---
output: github_document
---

<!-- README.md is generated from README.rmd. Please edit that file -->

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# stickylabeller

The `stickylabeller` package helps you label the facets in your ggplot2 plots. If you know how to use the [`glue`](https://cran.r-project.org/web/packages/glue/index.html) package, you know how to use `stickylabeller`!

## Installation

Install `stickylabeller` from GitHub using `devtools`:

```{r}
#| eval: false
devtools::install_github("jimjam-slam/stickylabeller")
```

## Use

The package has just one function: `label_glue`. Give it a string template to be processed by `glue`, and it'll return a labelling function that you can pass to `facet_*`:

```{r}
#| fig-cap: "facet_wrap labelled with two facet column values"

library(ggplot2)
library(stickylabeller)

# here's some example data: some random points in groups
mydf <- data.frame(
  x = 1:90,
  y = rnorm(90),
  red = rep(letters[1:3], 30),
  blue = c(rep(1, 30), rep(2, 30), rep(3, 30)))

# and here's a labelled plot!
ggplot(mydf) +
  geom_point(aes(x = x, y = y)) +
  facet_wrap(
    vars(red, blue),
    labeller = label_glue("Red is {red}\nand blue is {blue}"))
```

```{r}
#| fig-cap: "facet_grid labelled with one facet column value on each margin"

ggplot(mydf) +
  geom_point(aes(x = x, y = y)) +
  facet_grid(
    vars(red), vars(blue),
    labeller = label_glue(
      "Red is {red}",
      "Blue is {blue}"))
```

Your `label_glue` labeller can refer to any of the data frame columns included in the facetting formula. It can also use those columns in expressions, like:

```{r}
my_labeller <- label_glue("Red is {toupper(red)}\nand blue is {blue}")
```

### Numbering sequential facets

As well as the columns you include in the facetting specification, `stickylabeller` includes a few helper columns:

- `.n` numbers the facets numerically: `1`, `2`, `3`...
- `.l` numbers the facets using lowercase letters: `a`, `b`, `c`...
- `.L` numbers the facets using uppercase letters: `A`, `B`, `C`...
- `.r` numbers the facets using lowercase Roman numerals: `i`, `ii`, `iii`...
- `.R` numbers the facets using uppercase Roman numerals: `I`, `II`, `III`...

So you can automatically number your facets like:

```{r}
#| fig-cap: "facet_wrap labelled with sequential lowercase letters: (a), (b), (c) and so on"

ggplot(mydf) +
  geom_point(aes(x = x, y = y)) +
  facet_wrap(
    vars(red, blue),
    labeller = label_glue("({.l}) Red is {toupper(red)}\nand blue is {blue}"))
```

#### Limitations

* `.n`, `.l` and `.L` only work with `facet_wrap` for now. See [Issue #1](https://github.com/jimjam-slam/stickylabeller/issues/1).

### Including summary statistics in facet labels

As of stickylabeller 1.0, `label_glue` now accepts a `summary_data` argument. This accepts a data frame of pre-computed summary data for the facets; it should include the facet grouping columns, plus any other variables you'd like to reference in the templates.

For example:

```{r}
#| fig-cap: "facet_wrap labelled with two facet column values and two summary statistics, each numbered (a), (b), (c) and so on"
library(magrittr)
library(dplyr)

# summarise the data
mydf_stats = mydf %>%
  group_by(red, blue) %>%
  summarise(
    mean_y = sprintf("%#.2f", mean(y)),
    sd_y = sprintf("%#.2f", sd(y))) %>%
  ungroup()

# pass the summary data onto label_glue
ggplot(mydf) +
  geom_point(aes(x = x, y = y)) +
  facet_wrap(
    vars(red, blue),
    labeller = label_glue(
      "({.L}) Red = {red}, blue = {blue}\n(mean = {mean_y}, SD = {sd_y})",
      summary_data = mydf_stats))

```

Have fun! If you hit any snags, please feel free to [file an issue here](https://github.com/jimjam-slam/stickylabeller/issues) so that I can get on it! <3
